---
title: "Project_Random_Forests"
author: "Avneet Kaur"
date: "2022-12-05"
output: html_document
---

```{r}
df = read.csv("heart_failure_clinical_records_dataset.csv")
df$DEATH_EVENT  = as.factor(df$DEATH_EVENT)
attach(df)
```

Implementing Bagging considering DEATH_EVENT as the response variable and all other variables as predictor variables except time.
```{r}
library(randomForest)
set.seed(2331)
dfbag <- randomForest(DEATH_EVENT~., data=df, mtry=12, importance=TRUE)
dfbag
```

Importance plot
```{r, fig.width=10}
varImpPlot(dfbag)
```

Implementing Bagging on the the top 3 important predictor variables: "serum_creatinine","ejection_fraction","time"
```{r}
set.seed(2331)
dfbag1 <- randomForest(DEATH_EVENT~., data=df[,c("serum_creatinine","ejection_fraction","time")], mtry=3, importance=TRUE)
dfbag1
```

Implementing Random Forests

```{r}
set.seed(231)
dfbag2 <- randomForest(DEATH_EVENT~., data=df, importance=TRUE)
```
```{r}
dfbag2
```

Importance plots

```{r, fig.width=10}
varImpPlot(dfbag2)
```

Implementing Random forest on the the top 3 important predictor variables: "serum_creatinine","ejection_fraction","time"
```{r}
set.seed(233)
dfbag3 <- randomForest(DEATH_EVENT~., data=df[,c("serum_creatinine","ejection_fraction","time")], importance=TRUE)
```
```{r}
dfbag3
```