---
title: "Project"
author: "Avneet Kaur"
date: "2022-12-02"
output: html_document
---

```{r}
df = read.csv("heart_failure_clinical_records_dataset.csv")
head(df)
```

```{r}
attach(df)

glmfit <- glm(factor(DEATH_EVENT) ~. , data=df, family="binomial")
summary(glmfit)
```

On fitting logistic regression model with all the predictor variables, the summary output shows that "age", "ejection_fraction", "serum_creatinine", and "time" are the significant predictors and so we can remove the rest while fitting.

Applying Backward selection

```{r}
glmfit <- glm(factor(DEATH_EVENT) ~.-anaemia , data=df, family="binomial")
summary(glmfit)
```

```{r}
glmfit <- update(glmfit, .~. - smoking)
summary(glmfit)
```

```{r}
glmfit <- update(glmfit, .~. - high_blood_pressure)
summary(glmfit)
```

```{r}
glmfit <- update(glmfit, .~. - diabetes)
summary(glmfit)
```

```{r}
glmfit <- update(glmfit, .~. - platelets)
summary(glmfit)
```

```{r}
glmfit <- update(glmfit, .~. - creatinine_phosphokinase)
summary(glmfit)
```

```{r}
glmfit <- update(glmfit, .~. - sex)
summary(glmfit)
```

```{r}
glmfit <- update(glmfit, .~. - serum_sodium)
summary(glmfit)
```

```{r}
glmfit <- glm(factor(DEATH_EVENT)~age+ejection_fraction+serum_creatinine+time, data=df, family="binomial")
summary(glmfit)
```

We are now getting each of the predictor variables as significant.

Now let us try fitting logistic regression model by splitting the data into training and testing set.

1. Splitting the data into 80:20 ratio.
```{r}
st_num = 16756777
set.seed(st_num)                                                        
N = nrow(df) # total number of observations                                     
n = 0.8*N # number of observations in training                          
m = 0.2*N # number of observations in testing        

tr_ind = sample(N, n) # row index of training observations           
te_ind = setdiff(seq_len(N), tr_ind) # row index of test obs           

# Training Set                                                  
tr_set = df[tr_ind,] # training set
tr_x = tr_set[,-13] # stores the X inputs for the training set
tr_y = tr_set[,13] # stores the Y output for the training set

# Testing Set 
te_set = df[te_ind,] # test set 
te_x = te_set[,-13] # stores the X inputs for the test set
te_y = te_set[,13] # stores the Y output for the test set
```

Fitting the logistic regression model on the training set and then computing Misclassification rate by making prediction on the test set.

```{r}
glmfit <- glm(factor(DEATH_EVENT)~age+ejection_fraction+serum_creatinine+time, data=tr_set, family="binomial")
summary(glmfit)

te_pr = predict(glmfit, newdata = te_set, type = "response")

pred_class = rep(NA, length(te_pr))

for(i in 1:length(te_pr)){
  pred_class[i] = ifelse(te_pr[i] > 0.5, 1, 0)
}

tab = table(te_set$DEATH_EVENT, pred_class)
tab

# Misclassification rate
mcr <- (sum(tab)-sum(diag(tab)))/sum(tab)

```

```{r}
# Misclassification rate
mcr
```

Calculating other Performance evaluation metrics.

```{r, echo=FALSE}
library(MLmetrics)

# f1_Score = F1_Score(te_set$DEATH_EVENT, pred_class)
# f1_Score
# 
# logloss = LogLoss(pred_class, te_set$DEATH_EVENT)
# logloss
# 
# sensitivity <- Sensitivity(te_set$DEATH_EVENT, pred_class)
# sensitivity
# 
# recall <- Recall(te_set$DEATH_EVENT, pred_class)
# recall
# 
# precision <- Precision(te_set$DEATH_EVENT, pred_class)
# precision
# 
# specificity <- Specificity(te_set$DEATH_EVENT, pred_class)
# specificity

```

Now, let us try logistic regression using k-fold cross validation (k=5, 10).

```{r}
set.seed(9)
n = nrow(df)
k_list <- c(5,10)
msecv1 <- NA

for (j in 1:length(k_list)) {
  k = k_list[j]
  # reorder you observations
  rdf = df[sample(n),] 
  # Create k approx. equally size folds
  folds <- cut(1:n, breaks=k, labels=FALSE)

  for (i in 1:k) {
    train = which(folds != i)
    validation = which(folds == i)
    
    fit <- glm(factor(DEATH_EVENT)~age+ejection_fraction+serum_creatinine+time, data=df, subset=train, family="binomial")
    pred_class = ifelse(predict(fit, newdata=df[validation,], type="response")>0.5, 1, 0)
    tb = table(df[validation,]$DEATH_EVENT, pred_class)
    msecv1[i] = (sum(tb)-sum(diag(tb)))/sum(tb)
  }
  print(paste0("For k = ", k, " Cross-validation test error is ", sum(msecv1)/k, "."))
}
```

The cross-validated MSE estimate is smaller for k=10, so we will go forward with 10-fold cross-validation and evaluate model's performance.

```{r}
k = 10
msecv1 <- NA
# f1_Score1 <- NA
# logloss1 <- NA
# recall1 <- NA
# precision1 <- NA

folds <- cut(1:n, breaks=k, labels=FALSE)

for (i in 1:k) {
  train = which(folds != i)
  validation = which(folds == i)
    
  fit <- glm(factor(DEATH_EVENT)~age+ejection_fraction+serum_creatinine+time, data=df, subset=train, family="binomial")
  
  pred_class = ifelse(predict(fit, newdata=df[validation,], type="response")>0.5, 1, 0)
  
  tb = table(df[validation,]$DEATH_EVENT, pred_class)
  msecv1[i] = (sum(tb)-sum(diag(tb)))/sum(tb)
  # f1_Score1[i] = F1_Score(df[validation,]$DEATH_EVENT, pred_class)
  # 
  # logloss1[i] = LogLoss(pred_class, df[validation,]$DEATH_EVENT)
  # 
  # recall1[i] <- Recall(df[validation,]$DEATH_EVENT, pred_class)
  # 
  # precision1[i] <- Precision(df[validation,]$DEATH_EVENT, pred_class)

}

print(paste0("Cross-validation test error is ", sum(msecv1)/k, "."))
# print(paste0("Cross-validation F1 Score is ", sum(f1_Score1)/k, "."))
# print(paste0("Cross-validation LogLoss is ", sum(logloss1)/k, "."))
# print(paste0("Cross-validation Recall is ", sum(recall1)/k, "."))
# print(paste0("Cross-validation Precision is ", sum(precision1)/k, "."))

```

```{r}
k = 10
# msecv1 <- NA
# f1_Score1 <- NA
# logloss1 <- NA
# recall1 <- NA
# precision1 <- NA

folds <- cut(1:n, breaks=k, labels=FALSE)

for (i in 1:k) {
  train = which(folds != i)
  validation = which(folds == i)
    
  fit <- glm(factor(DEATH_EVENT)~age+ejection_fraction+serum_creatinine, data=df, subset=train, family="binomial")
  
  pred_class = ifelse(predict(fit, newdata=df[validation,], type="response")>0.5, 1, 0)
  
  tb = table(df[validation,]$DEATH_EVENT, pred_class)
  msecv1[i] = (sum(tb)-sum(diag(tb)))/sum(tb)
  # f1_Score1[i] = F1_Score(df[validation,]$DEATH_EVENT, pred_class)
  # 
  # logloss1[i] = LogLoss(pred_class, df[validation,]$DEATH_EVENT)
  # 
  # recall1[i] <- Recall(df[validation,]$DEATH_EVENT, pred_class)
  # 
  # precision1[i] <- Precision(df[validation,]$DEATH_EVENT, pred_class)

}

print(paste0("Cross-validation test error is ", sum(msecv1)/k, "."))
# print(paste0("Cross-validation F1 Score is ", sum(f1_Score1)/k, "."))
# print(paste0("Cross-validation LogLoss is ", sum(logloss1)/k, "."))
# print(paste0("Cross-validation Recall is ", sum(recall1)/k, "."))
# print(paste0("Cross-validation Precision is ", sum(precision1)/k, "."))
```